---
title: "Estimating Right-Wing Authoritarianism Values for United States Counties"
author: Mia Maines
subtitle: Using R Regression Modeling
date: today
date-format: long
---

# Introduction

The Far Right’s broad resurgence is probably one of the most noteworthy trends in the 21st century’s global political landscape, and politicians representing this shift to the right have gained prominence in not only the US, but in states such as Brazil, India, Russia, Turkey, Hungary, Italy, and Japan. It remains increasingly clear that this trend may be semi-permanent, with over a decade of increasing support for such movements among individuals of all walks of life. “Right wing authoritarianism” is a term not only describing a political identity but a set of beliefs and morals potentially bridging the gap between political science and psychology. A “right wing authoritarianism” scale was developed by psychologist Bob Altemeyer in 1981, describing a set of tendencies that might make one more predispositioned to supporting a right-wing authoritarian leader. These tendencies include high levels of submission to established authorities, willingness to engage in retaliatory actions to protect such authorities, and high degrees of conventionalism and traditionalism, whatever that may be in one’s culture. I will attempt to map such values in the United States, and the county-by-county spatial variance between them. I will take state-by-state or metro-by-metro survey data corresponding with right-wing authoritarian tendencies as defined by Altemeyer, and use a regression model to predict what socioeconomic traits could suggest right-wing authoritarianism in geographic regions.

Perhaps class is a primary predictor of RWA, or urban-rural location, or perhaps Pentecostalism. Modeling RWA by county will allow me to see the factors that correlate the most. 

# Materials and methods

Firstly, we examine existing survey data pertaining to right-wing authoritarianism in the United States. There are two primary surveyors I've found to have information pertaining to RWA-adjacent values: Gallup and the Public Religion Research Institute.
I take state-by-state and metro-by-metro information on RWA from these sources, and model them from Census Data pertaining to socioeconomic factors, as well as religion data from each county.

Code: The code associated with the project is well organized and easy to follow. Demonstrates mastery of R graphics and functions.

Data: The underlying data are publicly accessible via the web and downloaded/accessed within the Rmd script. If you want to use your own data, you must make it available on a website (e.g. Figshare) so that others are able to re-run your code.

You can do bullets like this:

-   The first most important thing
-   The second most important thing
-   The third most important thing

You can do numbers like this:

1.  The first most important thing
2.  The second most important thing
3.  The third most important thing

See <http://rmarkdown.rstudio.com/> for all the amazing things you can do.

Here's my first code chunk.

```{r}
x=3+4
```

Refer to output in your narrative like this: x=`r x` .

Load any required packages in a code chunk (you may need to install some packages):

```{r, message=F, warning=F}
library(tidyverse)
library(leaflet)
library(kableExtra)
library(htmlwidgets)
library(widgetframe)
knitr::opts_chunk$set(widgetframe_widgets_dir = 'widgets' ) 
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```

## Download and clean all required data

```{r}
n=20
data=data.frame(x=runif(n,-180,180),
                y=runif(n,-60,60),
                size = runif(n, 5, 20),
                category = factor(
                  sample(letters[1:5], n, replace = TRUE)
                  ),
                value = rnorm(n))
```

```{r, results='asis'}
data %>% 
  slice(1:10) %>% #show only 1:n rows
  kable(digits=2,align="c")%>% #make table and round to two digits
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive")) #apply other formatting
```

Add any additional processing steps here.

# Results

\[\~200 words\]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r, fig.width=6, fig.height=3, fig.cap="Map of completely random data"}
m <- leaflet(data) %>% 
  addTiles() %>% 
  addCircleMarkers(~x, ~y, radius = ~size,color = ~as.factor(category)) %>% 
  addPopups(~x[2], ~y[2], "Random popup")
m  # a map with the default OSM tile layer
```

```{r}
data %>% 
  ggplot(aes(x=x,y=y,col=category))+
  geom_point()
```

### Dygraphs Example

```{r}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures") |> 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01")) 
```

# Conclusions

\[\~200 words\]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

All sources are cited in a consistent manner
